---
title: |
  | YAPS - Yet Another Positioning Solver
  | OTN Telemetry Workshop Series, Dalhousie University, Halifax, Canada


output: github_document

---


[![][yaps_logo]](https://github.com/baktoft/yaps)   ![][otn_logo]   
  

***

Please make sure that these packages are installed.

```{r eval=FALSE}
install.packges(c('data.table',
                  'devtools', 
                  'dplyr',
                  'sp',
                  'leaflet',
                  'lubridate',
                  'ggplot2',
                  'caTools',
                  'viridis')
```

Please take a look at the TMB pages for info on installing TMB: https://github.com/kaskr/adcomp/wiki/Download
Most often, this line works `install.packages("TMB", type = "source")`

Then install yaps from github, load it and check it is working.
For this workshop, make sure to grab the 'dev_ows' branch i.e. `ref='dev_ows'`
```{r eval=FALSE}
devtools::install_github('baktoft/yaps', ref='dev_ows')
library(yaps)
testYaps()

```

If the last line returned a plot of a simple track with overlapping black and red lines, everything should be working.  

***  
The code for this part of the workshop is available in two formats - the code should be identical. All relevant files are found in the `rmds` folder. 

* ye good olde R-script
  - `rmds/otn_part1_setupAndTest.R`
  - `rmds/otn_part2_crayfish.R`
  - `rmds/otn_part3_hald.R`
* fancy-pants R-notebooks
  - `rmds/otn_part1_setupAndTest.nb.html`
  - `rmds/otn_part2_crayfish.nb.html`
  - `rmds/otn_part3_hald.nb.html`


***  


You are encouraged to take a look at the yaps readme: https://github.com/baktoft/yaps


If you can't get enough of YAPS have a look at the original paper:  
[Positioning of aquatic animals based on time-of-arrival and random walk models using YAPS (Yet Another Positioning Solver)](https://www.nature.com/articles/s41598-017-14278-z.pdf)  


And the recent pre-print including a step-by-step guide:
[Opening the black box of fish tracking using acoustic telemetry](https://www.biorxiv.org/content/10.1101/2019.12.16.877688v1)  

To keep updated, make sure to follow YAPS on [github](https://github.com/baktoft/yaps) and [researchgate](https://www.researchgate.net/project/YAPS-open-source-high-resolution-fish-tracking)  

## Sorry for the upload to github being a bit late!  



[yaps_logo]: rmds/yaps_logo_hex_100px.png
[otn_logo]: rmds/otn_logo.png

